Distributed Topology Discovery in Self-Assembled Nano Networks

### 2014-02-18 trunk/PAPER/disr.tex
Fixes

\label{sec:disr_concepts}
The main concept behind the kind of turn prohibition we want to
achieve is the \emph{segment}. A segment $S_y$ is basically a path of consecutive
nodes and links, starting with a link attached to a node belonging to
a different segment $S_x$ and ending with a link attached to a
node of another segment $S_z$. In other words, a segment
a path between connecting two other different segments. The example
in Figure~\ref{fig:segments} shows the segment with id $1.4$, starting
with a link connected to the node $9$, the going trought nodes
$4$,$3$,$2$ and ending with the link connected to the node $1$. Other
segments depicted are $9.2$ (link,node $10$,link) and $9.3$
(link from $9$ ,$8$,link,$7$,link,$6$,link,$5$,link to the $1$). Note
also that start/ending link can coincide and a segment can also be
formed by one single link (e.g. in $7.2$ or $5.4$).
An exception to the rule that a segment must connect two other
segments is the first segment established in the network, called
\emph{starting segment} which is a loop beginning/ending on the same
node.

@@

A detailed description of the \disr{} execution model as depicted in
the Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution of can be described as the execution
of the following phases: 
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node} 
\item Broadcasting and confirmation to create the first segment of the subnet
\item Parallel requests starting from assigned node to discover other
segments. 
\item Turn proibition in each confirmed segment.
\item Draining of remaining packet due timeouts
\end{enumerate}

@@

We distinct between two different kind of data stored at each node:
\emph{Dynamic Behaviour Status} (\emph{DBS}) and \emph{Local
Environment Data} (\emph{LED}).

The \emph{DBS} stores information in order to capture the dynamic
behaviour of the node  in the context of the whole
\disr{} execution phased. The \emph{DBS} can have the following values:

@@

The \emph{LED} are like a instanst snapshot of the \disr{} algorithm at each node,
consisting in the following variables:
\begin{itemize}
\item{\emph{segID}}:a value used to specify the segment to which the
node has been assigned or is candidate for being assigned.
\item{\emph{visited}}: a boolean value. When \emph{true}
then a \emph{segID} different from \texttt{NULL} specifies the segment 
to which the node has been assigned. 
\item{\emph{tvisited}}: a boolean value. If \emph{true}, the node is
being considered as candidate for a segment, and the \emph{segID} value
specifies the segment id for which the node is candidate. 
\item{\emph{link\_visited[]}}: an array
of values representing information about attached links, that is, the
segID of the segment owning each link. When \texttt{NULL}, the corresponding link has not yet been
assigned.
\item{\emph{link\_tvisited[]}}: an array of
values representing information about attached links, that is, the \emph{segID} of
the segment for which the link is candidate. When \texttt{NULL}, the link is
not candidate for any segment, i.e. maybe already assigned
or free.
\end{itemize}

@@

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[width=0.23\textwidth]{pictures/seq01.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq02.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq03.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq04.eps} \\
(a) & (b) & (c) & (d) \\
\includegraphics[width=0.23\textwidth]{pictures/seq05.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq06.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq07.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq08.eps} \\
(e) & (f) & (g) & (h) \\
\end{tabular}
\caption{Example of \disr{} segment setup: Nodes status are labeled as CandidateStarting($Cs$), Candidate ($C$), Free ($F$), Assigned ($A$) and Boostrap ($B$)}
\label{fig:disr_events}
\end{figure*}

@@

\item{\textbf{Bootnode Broadcasting}}: Each node receiving a \texttt{STARTING\_SEGMENT\_REQUEST},
when $Free$, forwards it to its free links using a flooding mechanism and
becoming $CandidateStarting$ $(g)$.
Each of its free links is then marked as \emph{tvisited} with the segment id
associated to the request. Note that a node that has already received a
\texttt{STARTING\_SEGMENT\_REQUEST} packet can simply ignore further packets
associated to the same request, having already contributed to the
flooding (Figure~\ref{fig:disr_events}-b).
%Note also that each of these packets has a max\_segment\_hops
%field to prevent packets from travelling undefinitively.

@@

\item{\textbf{Injecting Segment Requests}}: Each node in the
\emph{Assigned} status can initiate a search for a segment, by sending
a \texttt{SEGMENT\_REQUEST} across one of its free links $(c)$ (see
Figure~\ref{fig:disr_events}-e). 
Note that each \emph{Assigned} note can inject the request just after
becoming assigned, i.e. it is not needed that the whole confirmation
process in completed (Figure~\ref{disr_events}-f).
Note also that nodes previosly set as \emph{CandidateStarting}, when receiving 
a \texttt{SEGMENT\_REQUEST} can simply cancel their
status and set as $Candidate$ for that request, since this new
request means that the first starting segment has already been found $(h)$. 
\item{\textbf{Segment Confirmation}}: The find process is successful
when an already \emph{Assigned} node receives the
\texttt{SEGMENT\_REQUEST} packet. Then, a \texttt{SEGMENT\_CONFIRM} is
sent back along the path that originated the request
(Figure~\ref{fig:disr_events}-g), while the
node remain $Assigned$ (since it could confirm more segments). Each node
previously set as $Candidate$ for that segment id, when receiving the
a confirmation packet changes its status to $Assigned$ $(m)$ and forwards back
the same confirmation until it reaches the initiator of the request,
which then changes from $ActiveSearching$ to $Assigned$ $(d)$.

@@

\section{Detailed Node behaviour Model}
\label{sec:execution_model}

@@

The \emph{LED} variables at each node
($segID$,$visited$,$tvisited$,\emph{link\_visited[]},\emph{link\_tvisited[]})
are initially set as \texttt{NULL}, so the node and each of its links
are considered as free.
The only node that makes exception is the bootstrap node, initialized
in the \emph{Bootstrap} status. All the subsequent events happening at
each node are the consequence of its current dynamic status (DBS), the
packets received and the internal status represented by the
\emph{LED}. Note that in the following, when using expression like
``has a free link'' or ``free node'' we will assume the relationship beetween 
$LED$ and $DBS$ discussed in Section~\ref{sec:disr_data}. Further, for
each packet type all the cases not explicitly listed can simply be
ignored since are associated to invalid/inconsistent events of the
\disr{} execution model. A complete list of these cases can be checked
in the environment simulator source code at~\ref{nanoxim}.

@@

\subsection{Receiving a \texttt{SEGMENT\_CONFIRM}}
When the node status is \emph{Candidate} with a \emph{segID} corresponding to the one indicated in
the packet: the node set itself as $Assigned$ to the segment $SegID$. Further, it should forward this packet to the link where the
original \texttt{SEGMENT\_REQUEST} packet came from, so that all candidate nodes
can learn the segment id they belong to. \emph{LED} shoud be
updated from \emph{tvisited} to \emph{visited}.

@@

\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
\emph{Assigned} status can start a find segment process by injecting a
\texttt{SEGMENT\_REQUEST}. However, a more accurate decision when
defining \disr{} should be made in order to decide whether or not the
node must actually perform this action.  The point above is is
strictly related to the general question: which level of “parallelism”
should we allow in \disr{} ? The adopted approach is that, although
the nature of \disr{} itself is intrinsicly parallel, the use of
parallelism makes things work in a more complex way. In other words,
\disr{} is parallel when needed, but it does not exploit parallelism
as an “improving feature”. When not needed, things should be
serialized. For example: when a node is “ready searching” could start
several find segment process associated to the same segment, one for
each free link. But serializing this by investigating the free links
in order could be a simplest solution. We refer to this saying that we
avoid intra-node parallelism.  Note that while intra-node parallelism
can be avoided, trying to avoid inter-node parallelism could be more
complex than allowing the parallelism itself. Imagine for example the
effort of trying to coordinate nodes so that an unique finding node
process is actually running in the whole subnet. Thus, in contrast to
the intra-node parallelism, the inter-node parallelism is a structural
property of the \disr{} algorithm and should not be avoided.
%------------------------------------------------------------------------------

>>>>

\label{sec:disr_concepts}
The main concept behind the kind of turn prohibition we want to
achieve is the \emph{segment}. A segment $S_y$ is basically a path of consecutive
nodes and links, starting with a link attached to a node belonging to
a different segment $S_x$ and ending with a link attached to a
node of another segment $S_z$. In other words, a segment
a path between connecting two other different segments. The example
in Figure~\ref{fig:segments} shows the segment with id $1.4$, starting
with a link connected to the node $9$, the going through nodes
$4$,$3$,$2$ and ending with the link connected to the node $1$. Other
segments depicted are $9.2$ (link,node $10$,link) and $9.3$
(link from $9$ ,$8$,link,$7$,link,$6$,link,$5$,link to the $1$). Note
also that start/ending link can coincide and a segment can also be
formed by one single link (e.g. in $7.2$ or $5.4$).
An exception to the rule that a segment must connect two other
segments is the first segment established in the network, called
\emph{starting segment} which is a loop beginning/ending on the same
node.

@@

A detailed description of the \disr{} execution model as depicted in
the Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution of can be described as the execution
of the following phases: 
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node} 
\item Broadcasting and confirmation to create the first segment of the subnet
\item Parallel requests starting from assigned node to discover other
segments. 
\item Turn prohibition in each confirmed segment.
\item Draining of remaining packet due timeouts
\end{enumerate}

@@

We distinct between two different kind of data stored at each node:
\emph{Dynamic Behavior Status} (\emph{DBS}) and \emph{Local
Environment Data} (\emph{LED}).

The \emph{DBS} stores information in order to capture the dynamic
behavior of the node  in the context of the whole
\disr{} execution phased. The \emph{DBS} can have the following values:

@@

The \emph{LED} are like an instant snapshot of the \disr{} algorithm at each node,
consisting in the following variables:
\begin{itemize}
\item{\emph{segID}}:a value used to specify the segment to which the
node has been assigned or is candidate for being assigned.
\item{\emph{visited}}: a boolean value. When \emph{true}
then a \emph{segID} different from \texttt{NULL} specifies the segment 
to which the node has been assigned. 
\item{\emph{tvisited}}: a boolean value. If \emph{true}, the node is
being considered as candidate for a segment, and the \emph{segID} value
specifies the segment id for which the node is candidate. 
\item{\emph{link\_visited[]}}: an array
of values representing information about attached links, that is, the
segID of the segment owning each link. When \texttt{NULL}, the corresponding link has not yet been
assigned.
\item{\emph{link\_tvisited[]}}: an array of
values representing information about attached links, that is, the \emph{segID} of
the segment for which the link is candidate. When \texttt{NULL}, the link is
not candidate for any segment, i.e. maybe already assigned
or free.
\end{itemize}

@@

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[width=0.23\textwidth]{pictures/seq01.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq02.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq03.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq04.eps} \\
(a) & (b) & (c) & (d) \\
\includegraphics[width=0.23\textwidth]{pictures/seq05.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq06.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq07.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq08.eps} \\
(e) & (f) & (g) & (h) \\
\end{tabular}
\caption{Example of \disr{} segment setup: Nodes status are labeled as CandidateStarting($Cs$), Candidate ($C$), Free ($F$), Assigned ($A$) and Bootstrap ($B$)}
\label{fig:disr_events}
\end{figure*}

@@

\item{\textbf{Bootnode Broadcasting}}: Each node receiving a \texttt{STARTING\_SEGMENT\_REQUEST},
when $Free$, forwards it to its free links using a flooding mechanism and
becoming $CandidateStarting$ $(g)$.
Each of its free links is then marked as \emph{tvisited} with the segment id
associated to the request. Note that a node that has already received a
\texttt{STARTING\_SEGMENT\_REQUEST} packet can simply ignore further packets
associated to the same request, having already contributed to the
flooding (Figure~\ref{fig:disr_events}-b).
%Note also that each of these packets has a max\_segment\_hops
%field to prevent packets from traveling undefinitively.

@@

\item{\textbf{Injecting Segment Requests}}: Each node in the
\emph{Assigned} status can potentially initiate a new segment request.
So it check its links and if a free link is found it changes its
status to $ActiveSearching$ $(c)$, then sending
a \texttt{SEGMENT\_REQUEST} across the link $(c)$ (see
Figure~\ref{fig:disr_events}-e). 
Note that each \emph{Assigned} node can inject the request just after
becoming assigned, i.e. it is not needed that the whole confirmation
process in completed (Figure~\ref{fig:disr_events}-f).
Note also that nodes previously set as \emph{CandidateStarting}, when receiving 
a \texttt{SEGMENT\_REQUEST} can simply cancel their
status and set as $Candidate$ for that request, since this new
request means that the first starting segment has already been found $(h)$. 
\item{\textbf{Segment Confirmation}}: The find process is successful
when an already \emph{Assigned} node receives the
\texttt{SEGMENT\_REQUEST} packet. Then, a \texttt{SEGMENT\_CONFIRM} is
sent back along the path that originated the request
(Figure~\ref{fig:disr_events}-g), while the
node remain $Assigned$ (since it could confirm more segments). Each node
previously set as $Candidate$ for that segment id, when receiving the
a confirmation packet changes its status to $Assigned$ $(m)$ and forwards back
the same confirmation until it reaches the initiator of the request,
which then changes from $ActiveSearching$ to $Assigned$ $(d)$.

@@

\section{Detailed Node behavior Model}
\label{sec:execution_model}

@@

The \emph{LED} variables at each node
($segID$,$visited$,$tvisited$,\emph{link\_visited[]},\emph{link\_tvisited[]})
are initially set as \texttt{NULL}, so the node and each of its links
are considered as free.
The only node that makes exception is the bootstrap node, initialized
in the \emph{Bootstrap} status. All the subsequent events happening at
each node are the consequence of its current dynamic status (DBS), the
packets received and the internal status represented by the
\emph{LED}. Note that in the following, when using expression like
``has a free link'' or ``free node'' we will assume the relationship between 
$LED$ and $DBS$ discussed in Section~\ref{sec:disr_data}. Further, for
each packet type all the cases not explicitly listed can simply be
ignored since are associated to invalid/inconsistent events of the
\disr{} execution model. A complete list of these cases can be checked
in the environment simulator source code at~\ref{nanoxim}.

@@

\subsection{Receiving a \texttt{SEGMENT\_CONFIRM}}
When the node status is \emph{Candidate} with a \emph{segID} corresponding to the one indicated in
the packet: the node set itself as $Assigned$ to the segment $SegID$. Further, it should forward this packet to the link where the
original \texttt{SEGMENT\_REQUEST} packet came from, so that all candidate nodes
can learn the segment id they belong to. \emph{LED} should be
updated from \emph{tvisited} to \emph{visited}.

@@

\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
\emph{Assigned} status can start a find segment process by injecting a
\texttt{SEGMENT\_REQUEST}. However, a more accurate decision when
defining \disr{} should be made in order to decide whether or not the
node must actually perform this action.  The point above is is
strictly related to the general question: which level of “parallelism”
should we allow in \disr{} ? The adopted approach is that, although
the nature of \disr{} itself is intrinsically parallel, the use of
parallelism makes things work in a more complex way. In other words,
\disr{} is parallel when needed, but it does not exploit parallelism
as an “improving feature”. When not needed, things should be
serialized. For example: when a node is “ready searching” could start
several find segment process associated to the same segment, one for
each free link. But serializing this by investigating the free links
in order could be a simplest solution. We refer to this saying that we
avoid intra-node parallelism.  Note that while intra-node parallelism
can be avoided, trying to avoid inter-node parallelism could be more
complex than allowing the parallelism itself. Imagine for example the
effort of trying to coordinate nodes so that an unique finding node
process is actually running in the whole subnet. Thus, in contrast to
the intra-node parallelism, the inter-node parallelism is a structural
property of the \disr{} algorithm and should not be avoided.
%------------------------------------------------------------------------------

### 2014-02-18 trunk/PAPER/paper.tex
Fixed style
Fixes

% Elsevier specific styles
%------------------------------------------------------------------------------
\documentclass[a4]{article}
%\documentclass[review]{elsarticle}
\usepackage{lineno,hyperref}

@@

\begin{abstract}
In this paper we present \disr{}, a distributed approach to topology
discovery and defect mapping in a self-assembled nano
networks-on-chip. The main aim is achieving the already-proven
properties of segment-based deadlock freedom without neither requiring
a topology graph as input, nor needing a centralized algorithm to
configure network paths.  After introducing the conceptual elements
and the execution model of \disr{}, we show how the opensource Nanoxim
platform has been used to evaluate the proposed approach in the
process of discovering irregular network topology while establishing
network segments. Comparison against a tree-based approach show how
\disr{} still preserves some important properties (coverage, defect
tolerance, scalability) while avoiding resource hungry solutions such
as virtual channels and hardware redundancy. Finally, we propose a
gate-level hardware implementation of the required control logic and
storage for \disr{}, demonstrating a relatively acceptable impact
ranging from 10 to about 20\% of the budget of transistors available
for each node.
\end{abstract}

>>>>

% Elsevier specific styles
%------------------------------------------------------------------------------
\documentclass[preprint]{elsarticle}
%\documentclass[3p]{elsarticle}
\usepackage{lineno,hyperref}

@@

\begin{abstract}
In this paper we present \disr{}, a distributed approach to topology
discovery and defect mapping in a self-assembled nano
network-on-chip. The main aim is achieving the already-proven
properties of segment-based deadlock freedom without neither requiring
a topology graph as input, nor needing a centralized algorithm to
configure network paths.  After introducing the conceptual elements
and the execution model of \disr{}, we show how the open-source Nanoxim
platform has been used to evaluate the proposed approach in the
process of discovering irregular network topology while establishing
network segments. Comparison against a tree-based approach shows how
\disr{} still preserves some important properties (coverage, defect
tolerance, scalability) while avoiding resource hungry solutions such
as virtual channels and hardware redundancy. Finally, we propose a
gate-level hardware implementation of the required control logic and
storage for \disr{}, demonstrating a relatively acceptable impact
ranging from 10 to about 20\% of the budget of transistors available
for each node.
\end{abstract}
